---
title: "R Notebook"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    keep_md: yes
  pdf_document: default
---

```{r}
# CHANGE PATH HERE:

mypath = "D:\\_nj_\\-github-\\Resources\\2022-07-21\\Evolution\\";


# YOU need to install these packages for this to work.
library(SentimentAnalysis);
library(wordcloud2);
library(wordcloud);
library(tm);
library(textclean);
library(slam);
```

# Inputs

I am comparing the ORIG 1828 NOAH WEBSTER AMERICAN DICTIONARY to the COPY of 1844.


## [1828] ORIG

```{r}

myFile = paste0(mypath,"1828.txt");
#ORIG = readChar(myFile, file.info(myFile)$size);
ORIG = readLines(myFile);
length(ORIG);
ORIG = unique(ORIG);  # replicates are same word with "multiple meanings" ... still ? SPELLED THE SAME
length(ORIG);
```



## [1844] COPY

```{r}

myFile = paste0(mypath,"1844.txt");
#COPY = readChar(myFile, file.info(myFile)$size);
COPY = readLines(myFile);
length(COPY);
COPY = unique(COPY);  # replicates are same word with "multiple meanings" ... still ? SPELLED THE SAME
length(COPY);
```

```{r}
ALL = union(ORIG,COPY);
```


## Summary

```{r}
if(!file.exists(paste0(mypath,"BOTH.txt")))
{
ALL = union(ORIG,COPY);
OLD = setdiff(ORIG,COPY);
NEW = setdiff(COPY,ORIG);
BOTH = intersect(ORIG,COPY);

writeLines(sort(ALL), paste0(mypath,"ALL.txt"));
writeLines(sort(OLD), paste0(mypath,"OLD.txt"));
writeLines(sort(NEW), paste0(mypath,"NEW.txt"));
writeLines(sort(BOTH), paste0(mypath,"BOTH.txt"));
}
```

```{r}
if(!file.exists(paste0(mypath,"BOTH.txt")))
{
df = as.data.frame(ALL);
colnames(df) = "words";

df$ALL = 1;
df$OLD = as.integer( is.element(ALL,OLD) );
df$NEW = as.integer( is.element(ALL,NEW) );
df$BOTH = as.integer( is.element(ALL,BOTH) );

rownames(df) = ALL;
df;
}
```

```{r}
# FULL
# STARTS
# END

revS = function(str, sep="", delimiter="")
{
  vec = toVec(str,sep);
  vecR = rev(vec);
  toStr(vecR, delimiter);
}

toStr = function(vec, delimiter="")
{
paste0(vec, collapse = delimiter);  
}
toVec = function(str, sep="")
{
  strsplit(str, sep, fixed=TRUE)[[1]];
}

# 1,2,3,4,5
# 26, 26^2, 26^3, 26^4, 26^5

parseWord = function(word)
{
  mylist = list();
  mylist$word = word;
  
  wl = nchar(word);
    mylist$wl = wl;
    
  
  start = end = full = NULL;
  
  ## let's do start and end ...
  myl = 5;
  if(wl < myl) { myl = wl; }
  
  
  stem = substr(word, 1, myl);
    now = substring(stem, 1, 1:myl);
    
    start = c(start, now);
    full = c(full, now);
    
    stemR = substr(revS(word), 1, myl);
    nowR = substring(stemR, 1, 1:myl);
    
    end = c(end, unlist(lapply(nowR, FUN=revS)) );
  
  ## let's now step into the middle of the word ... ONLY do full
  n = 1;
  myl = 5;
  # loops = wl - myl;
  loops = wl;
  
  if(loops > 0)
  {
  for(n in 1:loops)
  {
  
  stem = substr(word, n+1, myl+n);
    now = unique(substring(stem, 1, 1:myl));
    full = c(full, now);
  }
  }
  
  full[nchar(full)>0] 
  
  mylist$full = full[nchar(full)>0] ;
  mylist$start = start[nchar(start)>0] ;
  mylist$end = end[nchar(end)>0] ;
  
  mylist;
  
  
  
  
}


parseWord("abatis");
parseWord("recommissioning");
parseWord("a");
parseWord("the");
parseWord("apply");
```

## This will take some time, so let's cache the result ...

```{r}
myFile = paste0(mypath,"raw-ch.RDS");
timer.start = as.numeric( Sys.time() );
if(!file.exists(myFile))
{
  res = list();
  for(i in 1:length(ALL))
  {
    word = ALL[i];
    res[[word]] = parseWord(word);
    
    if(i %% 10000 == 0)
    {
    print(paste0(i," : ",word));
    }
    
  }


## eend lop

saveRDS(res, file=myFile);
}

timer.end = as.numeric( Sys.time() );

timer.diff = timer.end - timer.start;
print(paste0( timer.diff, " seconds "));
```

```{r}
# grab from CACHE ...
parsed = invisible( readRDS(myFile) );
```


## Organize the features into the df

```{r}
myFile2 = paste0(mypath,"features-ch.RDS");
timer.start = as.numeric( Sys.time() );
if(!file.exists(myFile2))
{
  res = list();
    m = c("OLD","NEW","BOTH");
    d = c("full", "start", "end");
    ### INIT ###
  for(mm in 1:3) 
    { 
    whichM = m[mm];
    res[[whichM]] = list(); 
  for(j in 1:3) 
    { 
    whichD = d[j];
    res[[whichM]][[whichD]] = list(); 
    for(i in 1:5)
    {
      res[[whichM]][[whichD]][[i]] = list();
    }
  }
    
  }
    ## END INIT

  for(i in 1:length(ALL))
  {
    word = ALL[i];
    data = parsed[[word]];
    
    whichM = 0;
    if(is.element(word,OLD)) {whichM = 1;}
    if(is.element(word,NEW)) {whichM = 2;}
    if(is.element(word,BOTH)) {whichM = 3;}
    
    for(j in 1:3) 
    { 
    whichD = d[j];
    data.j = data[[whichD]];
    for(k in 1:length(data.j))
      {
      ss = data.j[k];
      sl = nchar(ss);
      
      if(is.null(res[[whichM]][[whichD]][[sl]][[ss]]))
      {
        res[[whichM]][[whichD]][[sl]][[ss]] = 0;
      }
      
      res[[whichM]][[whichD]][[sl]][[ss]] = 1 + res[[whichM]][[whichD]][[sl]][[ss]]; # increment
      
      # names(res[[whichD]][[sl]]);
      # we will alphabatize them ... 
      
      }
    
    }
    
    if(i %% 1000 == 0)
    {
    print(paste0(i," : ",word));
    }
    
  }




saveRDS(res, file=myFile2);
}

timer.end = as.numeric( Sys.time() );

timer.diff = timer.end - timer.start;
print(paste0( timer.diff, " seconds "));
```

```{r}
# grab from CACHE ...
counted = invisible( readRDS(myFile2) );
```

## BUILD DATAFRAME for FREQ table
```{r}

m = c("OLD","NEW","BOTH");
d = c("full", "start", "end");

myFile3 = paste0(mypath,"df-ch.RDS");
timer.start = as.numeric( Sys.time() );
if(!file.exists(myFile3))
{
  nnn = 0;
  # counted
  df = NULL;
  
  for(mm in 1:3) 
    { 
    whichM = m[mm]; 
  for(j in 1:3) 
    { 
    whichD = d[j];
    for(i in 1:5)
    {
      # res[[whichM]][[whichD]][[i]] = list();
      mynames = sort(names(counted[[whichM]][[whichD]][[i]]));
      
      nnl = length(mynames);
      print( paste0(whichM," => ",whichD," => ",i," => ",nnl," => ",nnn ) );
      for(nn in 1:nnl)
      {
        nnn = 1 + nnn;
        myname = mynames[nn];
        myvalue = counted[[whichM]][[whichD]][[i]][[myname]];
        
        #Mmyname = paste0(myname,"_");
        #if(whichD == "end") { Mmyname = paste0("_",myname); }
        #if(whichD == "full") { Mmyname = paste0("_",myname,"_"); }
        
      row = c(whichM, whichD, i, myname, myvalue);
      #row = c(whichM, whichD, i, myname, Mmyname, myvalue);
      df = rbind(df, row);
      
      if(nn %% 1000 == 0)
    {
    print(paste0("               ",nn," / ",nnl," : ",myname," --> ", myvalue));
      }
      
      }
      
    }
  }
    
  }
    ## END INIT
  
  

df = as.data.frame(df);
colnames(df) = c("SET","DIRECTION", "NUMBER", "STEM", "FREQ");

saveRDS(df, file=myFile3);  
}
  
  
timer.end = as.numeric( Sys.time() );

timer.diff = timer.end - timer.start;
print(paste0( timer.diff, " seconds "));




```


```{r}
FEATURES = readRDS(myFile3);

FEATURES$NUMBER = as.numeric(FEATURES$NUMBER);
FEATURES$FREQ = as.numeric(FEATURES$FREQ);

```

# ---FULL (one-length)
```{r}
# rows are features
# cols what we want to compare

# JUST FULL

SUB = subset(FEATURES, DIRECTION=="full");
SUB = subset(SUB, NUMBER==1);
SUB$STEM2 = paste0("_",SUB$STEM,"_");

SUB.BOTH = subset(SUB, SET=="BOTH");
SUB.OLD = subset(SUB, SET=="OLD");
SUB.NEW = subset(SUB, SET=="NEW");

FREQ.BOTH = SUB.BOTH[,c(6,5)];
FREQ.OLD = SUB.OLD[,c(6,5)];
FREQ.NEW = SUB.NEW[,c(6,5)];


TABLE.FREQ = merge(FREQ.BOTH, FREQ.OLD, by="STEM2", all.x=TRUE, all.y=TRUE);
TABLE.FREQ = merge(TABLE.FREQ, FREQ.NEW, by="STEM2", all.x=TRUE, all.y=TRUE);
colnames(TABLE.FREQ) = c("FEATURE", "BOTH", "OLD", "NEW");
rownames(TABLE.FREQ) = TABLE.FREQ$FEATURE;

TABLE.FREQ[is.na(TABLE.FREQ)] = 0;
TABLE.FREQ;


```

## SETUP X
```{r}
names.features = TABLE.FREQ$FEATURE;
names.docs = colnames(TABLE.FREQ)[-1];
X = t( TABLE.FREQ[,-1] );
rownames(X) = names.docs;
colnames(X) = names.features;
```

### HCLUST
```{r}
Xclust = hclust(dist(scale(t(X))));
# transpose, then scale?
plot(Xclust);
```

### Compute bigraph (BOTH,OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X));  
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```







### Compute bigraph (OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X[-1,]));  # GET RID OF BOTH
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```





## TF-IDF
```{r}
# https://www.mathworks.com/help/textanalytics/ref/bagofwords.tfidf.html

X.tf = colSums(X);  # term freq
X.idf = colSums(X != 0); # document freq

# NORM technique?
# https://en.wikipedia.org/wiki/Tf%E2%80%93idf#Inverse_document_frequency
X.tf.s = log(1 + X.tf);
X.idf.s = log( nrow(X) / (1+X.idf) ) + 1;

X.tf.idf.s = X.tf.s * X.idf.s;  # PAIRWISE products

# https://en.wikipedia.org/wiki/Latent_semantic_analysis


## scaled X
Xs = X * X.tf.idf.s # PAIRWISE products

```

### HCLUST
```{r}
Xclust = hclust(dist(scale(t(Xs))));
# transpose, then scale? No scale on Xs? double scale?
plot(Xclust);
```

### Compute bigraph (BOTH,OLD,NEW)
```{r}
# prcomp vs princomp?
Xs.PCA = prcomp(t(Xs));  
summary(Xs.PCA);
str(Xs.PCA);
Xs.PCA;
biplot(Xs.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(Xs.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(Xs.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```


### Compute bigraph (OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X[-1,]));  # GET RID OF BOTH
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```



# ---FULL (two-length)
```{r}
# rows are features
# cols what we want to compare

# JUST FULL

SUB = subset(FEATURES, DIRECTION=="full");
SUB = subset(SUB, NUMBER==2);
SUB$STEM2 = paste0("_",SUB$STEM,"_");

SUB.BOTH = subset(SUB, SET=="BOTH");
SUB.OLD = subset(SUB, SET=="OLD");
SUB.NEW = subset(SUB, SET=="NEW");

FREQ.BOTH = SUB.BOTH[,c(6,5)];
FREQ.OLD = SUB.OLD[,c(6,5)];
FREQ.NEW = SUB.NEW[,c(6,5)];


TABLE.FREQ = merge(FREQ.BOTH, FREQ.OLD, by="STEM2", all.x=TRUE, all.y=TRUE);
TABLE.FREQ = merge(TABLE.FREQ, FREQ.NEW, by="STEM2", all.x=TRUE, all.y=TRUE);
colnames(TABLE.FREQ) = c("FEATURE", "BOTH", "OLD", "NEW");
rownames(TABLE.FREQ) = TABLE.FREQ$FEATURE;

TABLE.FREQ[is.na(TABLE.FREQ)] = 0;
TABLE.FREQ;


```
## SETUP X
```{r}
names.features = TABLE.FREQ$FEATURE;
names.docs = colnames(TABLE.FREQ)[-1];
X = t( TABLE.FREQ[,-1] );
rownames(X) = names.docs;
colnames(X) = names.features;
```

### HCLUST
```{r}
Xclust = hclust(dist(scale(t(X))));
# transpose, then scale?
plot(Xclust);
```

### Compute bigraph  (BOTH,OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X));  
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```







### Compute bigraph (OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X[-1,]));  # GET RID OF BOTH
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```





## TF-IDF
```{r}
# https://www.mathworks.com/help/textanalytics/ref/bagofwords.tfidf.html

X.tf = colSums(X);  # term freq
X.idf = colSums(X != 0); # document freq

# NORM technique?
# https://en.wikipedia.org/wiki/Tf%E2%80%93idf#Inverse_document_frequency
X.tf.s = log(1 + X.tf);
X.idf.s = log( nrow(X) / (1+X.idf) ) + 1;

X.tf.idf.s = X.tf.s * X.idf.s;  # PAIRWISE products

# https://en.wikipedia.org/wiki/Latent_semantic_analysis


## scaled X
Xs = X * X.tf.idf.s # PAIRWISE products

```

### HCLUST
```{r}
Xclust = hclust(dist(scale(t(Xs))));
# transpose, then scale? No scale on Xs? double scale?
plot(Xclust);
```

### Compute bigraph (BOTH,OLD,NEW)
```{r}
# prcomp vs princomp?
Xs.PCA = prcomp(t(Xs));  
summary(Xs.PCA);
str(Xs.PCA);
Xs.PCA;
biplot(Xs.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(Xs.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(Xs.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```


### Compute bigraph (OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X[-1,]));  # GET RID OF BOTH
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```



# ---START (one-length)
```{r}
# rows are features
# cols what we want to compare

# JUST START

SUB = subset(FEATURES, DIRECTION=="start");
SUB = subset(SUB, NUMBER==1);
SUB$STEM2 = paste0(SUB$STEM,"_");

SUB.BOTH = subset(SUB, SET=="BOTH");
SUB.OLD = subset(SUB, SET=="OLD");
SUB.NEW = subset(SUB, SET=="NEW");

FREQ.BOTH = SUB.BOTH[,c(6,5)];
FREQ.OLD = SUB.OLD[,c(6,5)];
FREQ.NEW = SUB.NEW[,c(6,5)];


TABLE.FREQ = merge(FREQ.BOTH, FREQ.OLD, by="STEM2", all.x=TRUE, all.y=TRUE);
TABLE.FREQ = merge(TABLE.FREQ, FREQ.NEW, by="STEM2", all.x=TRUE, all.y=TRUE);
colnames(TABLE.FREQ) = c("FEATURE", "BOTH", "OLD", "NEW");
rownames(TABLE.FREQ) = TABLE.FREQ$FEATURE;

TABLE.FREQ[is.na(TABLE.FREQ)] = 0;
TABLE.FREQ;


```
## SETUP X
```{r}
names.features = TABLE.FREQ$FEATURE;
names.docs = colnames(TABLE.FREQ)[-1];
X = t( TABLE.FREQ[,-1] );
rownames(X) = names.docs;
colnames(X) = names.features;
```


### HCLUST
```{r}
Xclust = hclust(dist(scale(t(X))));
# transpose, then scale?
plot(Xclust);
```

### Compute bigraph (BOTH,OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X));  
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```







### Compute bigraph (OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X[-1,]));  # GET RID OF BOTH
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```





## TF-IDF
```{r}
# https://www.mathworks.com/help/textanalytics/ref/bagofwords.tfidf.html

X.tf = colSums(X);  # term freq
X.idf = colSums(X != 0); # document freq

# NORM technique?
# https://en.wikipedia.org/wiki/Tf%E2%80%93idf#Inverse_document_frequency
X.tf.s = log(1 + X.tf);
X.idf.s = log( nrow(X) / (1+X.idf) ) + 1;

X.tf.idf.s = X.tf.s * X.idf.s;  # PAIRWISE products

# https://en.wikipedia.org/wiki/Latent_semantic_analysis


## scaled X
Xs = X * X.tf.idf.s # PAIRWISE products

```

### HCLUST
```{r}
Xclust = hclust(dist(scale(t(Xs))));
# transpose, then scale? No scale on Xs? double scale?
plot(Xclust);
```

### Compute bigraph (BOTH,OLD,NEW)
```{r}
# prcomp vs princomp?
Xs.PCA = prcomp(t(Xs));  
summary(Xs.PCA);
str(Xs.PCA);
Xs.PCA;
biplot(Xs.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(Xs.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(Xs.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```


### Compute bigraph (OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X[-1,]));  # GET RID OF BOTH
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```



# ---START (two-length)
```{r}
# rows are features
# cols what we want to compare

# JUST START

SUB = subset(FEATURES, DIRECTION=="start");
SUB = subset(SUB, NUMBER==2);
SUB$STEM2 = paste0(SUB$STEM,"_");

SUB.BOTH = subset(SUB, SET=="BOTH");
SUB.OLD = subset(SUB, SET=="OLD");
SUB.NEW = subset(SUB, SET=="NEW");

FREQ.BOTH = SUB.BOTH[,c(6,5)];
FREQ.OLD = SUB.OLD[,c(6,5)];
FREQ.NEW = SUB.NEW[,c(6,5)];


TABLE.FREQ = merge(FREQ.BOTH, FREQ.OLD, by="STEM2", all.x=TRUE, all.y=TRUE);
TABLE.FREQ = merge(TABLE.FREQ, FREQ.NEW, by="STEM2", all.x=TRUE, all.y=TRUE);
colnames(TABLE.FREQ) = c("FEATURE", "BOTH", "OLD", "NEW");
rownames(TABLE.FREQ) = TABLE.FREQ$FEATURE;

TABLE.FREQ[is.na(TABLE.FREQ)] = 0;
TABLE.FREQ;


```
## SETUP X
```{r}
names.features = TABLE.FREQ$FEATURE;
names.docs = colnames(TABLE.FREQ)[-1];
X = t( TABLE.FREQ[,-1] );
rownames(X) = names.docs;
colnames(X) = names.features;
```


### HCLUST
```{r}
Xclust = hclust(dist(scale(t(X))));
# transpose, then scale?
plot(Xclust);
```

### Compute bigraph (BOTH,OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X));  
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```







### Compute bigraph (OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X[-1,]));  # GET RID OF BOTH
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```





## TF-IDF
```{r}
# https://www.mathworks.com/help/textanalytics/ref/bagofwords.tfidf.html

X.tf = colSums(X);  # term freq
X.idf = colSums(X != 0); # document freq

# NORM technique?
# https://en.wikipedia.org/wiki/Tf%E2%80%93idf#Inverse_document_frequency
X.tf.s = log(1 + X.tf);
X.idf.s = log( nrow(X) / (1+X.idf) ) + 1;

X.tf.idf.s = X.tf.s * X.idf.s;  # PAIRWISE products

# https://en.wikipedia.org/wiki/Latent_semantic_analysis


## scaled X
Xs = X * X.tf.idf.s # PAIRWISE products

```

### HCLUST
```{r}
Xclust = hclust(dist(scale(t(Xs))));
# transpose, then scale? No scale on Xs? double scale?
plot(Xclust);
```

### Compute bigraph (BOTH,OLD,NEW)
```{r}
# prcomp vs princomp?
Xs.PCA = prcomp(t(Xs));  
summary(Xs.PCA);
str(Xs.PCA);
Xs.PCA;
biplot(Xs.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(Xs.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(Xs.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```


### Compute bigraph (OLD,NEW)
```{r}
# prcomp vs princomp?
X.PCA = prcomp(t(X[-1,]));  # GET RID OF BOTH
summary(X.PCA);
str(X.PCA);
X.PCA;
biplot(X.PCA, 
       xlab=paste0("PC1 VAF: ",round(summary(X.PCA)$importance[2,1] * 100,1 ), "%"),
       ylab=paste0("PC2 VAF: ",round(summary(X.PCA)$importance[2,2] * 100,1 ), "%"),
      );
```



